### 2025-09-12 cluster-analysis.qmd
20250910 - machine learning additions/edits
20250911 - no longer errors out

There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, hierarchical clustering (aka connectivity-based clustering), etc.
An overview of approaches to cluster analysis in `R` is provided by @Kassambara2017a.
In this chapter, we focus on examples using model-based clustering with the `R` package `mclust` [@R-mclust; @Scrucca2023_packages], which uses Gaussian finite mixture modeling.
The various types of `mclust` models are provided here:
<https://mclust-org.github.io/mclust/reference/mclustModelNames.html>.

@@

The model-based bootstrap clustering of Running Backs' fantasy points is unable to run due to an error:

@@

Thus, we cannot use the following code, which would otherwise summarize the model results, specify the number of tiers, and plot model comparisons:

numTiersRB <- as.numeric(summary(tiersRB_boostrap)[,"Length"][1]) # or could specify the number of teams manually

>>>>

There are many approaches to cluster analysis, including model-based clustering, density-based clustering, centroid-based clustering, hierarchical clustering (aka connectivity-based clustering), etc.
In general, cluster analysis intends to maximize intracluster similarities and to minimize intercluster similarities [@Ramasubramanian2016].
An overview of approaches to cluster analysis in `R` is provided by @Kassambara2017a.
In this chapter, we focus on examples using model-based clustering with the `R` package `mclust` [@R-mclust; @Scrucca2023_packages], which uses Gaussian finite mixture modeling.
The various types of `mclust` models are provided here:
<https://mclust-org.github.io/mclust/reference/mclustModelNames.html>.

@@

<!--
The model-based bootstrap clustering of Running Backs' fantasy points is unable to run due to an error:
-->

@@

<!--
Thus, we cannot use the following code, which would otherwise summarize the model results, specify the number of tiers, and plot model comparisons:
-->

```{r}
numTiersRB <- as.numeric(summary(tiersRB_boostrap)[,"Length"][1]) # or could specify the number of teams manually
